---
title: "Covid"
author: "Pinyu Chen, Yaning Wu"
date: "12/1/2021"
output:
  pdf_document:
    latex_engine: xelatex
---
## Import data
```{r}
library(tidyverse)
OCD <- read_csv("../data/owid-covid-data.csv")
head(OCD)

INCOME <- read_csv("../data/Metadata_Country_API_4_DS2_en_csv_v2_3160069.csv")
colnames(INCOME) <- c("Country_Code", "Region", "IncomeGroup")
glimpse(INCOME)
```
## Check the number of NA
```{r}
OCD %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = c(`iso_code`:`excess_mortality_cumulative_per_million`),
               names_to = "colnames", 
               values_to = "value_NA") %>%
  arrange(desc(value_NA)) %>%
  filter(value_NA > 60000)
```

## Select relative columns and deleted rows which population is NA
```{r}
OCD %>%
  select(iso_code, continent,location,date,total_cases,total_deaths,population,population_density,median_age,gdp_per_capita,diabetes_prevalence,human_development_index) -> OCD1

glimpse(OCD1)

OCD1 %>% 
  filter(!is.na(population))->OCD1

OCD1 %>%
  mutate(iso_code = parse_factor(iso_code),
         continent = parse_factor(continent),
         location = parse_factor(location)) -> OCD2
glimpse(OCD2)

nlevels(OCD2$location)
```

## Filter out the NA from INCOME data set
```{r}
INCOME %>%
  select(Country_Code, IncomeGroup) %>%
  filter(!is.na(IncomeGroup)) %>%
  mutate(IncomeGroup = str_replace_all(IncomeGroup, "\\s", "_"),
         IncomeGroup = parse_factor(IncomeGroup),
         Country_Code = parse_factor(Country_Code))-> INCOME1
glimpse(INCOME1)

levels(INCOME1$IncomeGroup)
```

## Combine INCOME data set to OCD data set
```{r}
OCD2 %>%
  left_join(INCOME1, by = c("iso_code" = "Country_Code")) %>%
  filter(!is.na(IncomeGroup))-> OCD3
OCD3
```

## Create proportion of total cases and deaths for each country
```{r}
OCD3 %>%
  group_by(iso_code, continent, location) %>%
  filter(date == max(date)) %>%
  filter(date != "2021-12-01") # this check the location that do not have 2021-12-01's record

OCD3 %>% 
  filter(date==max(date))->OCD4 #OCD4 is the table that only 2021-12-01 per country

OCD4 %>% 
  mutate(cases_per_person = total_cases/population,
         deaths_per_person = total_deaths/population) %>% 
  filter((!is.na(cases_per_person)) & (!is.na(deaths_per_person)))->OCD4  # This step is for removing proportion which has na

OCD4 %>% 
  slice_sample(n=30)

OCD4 %>%
  filter(across(everything(), ~ !is.na(.x))) -> Test.OCD4
```

## Adjust level of IncomeGroup
```{r}
OCD4 %>%
  mutate(IncomeGroup = fct_relevel(IncomeGroup, "Low_income", after = nlevels(IncomeGroup)),
         IncomeGroup = fct_relevel(IncomeGroup, "Lower_middle_income", after = 2L)) -> OCD4
```

## Mean cases/deaths per person vs date
```{r}
OCD3 %>% 
  mutate(cases_per_person = total_cases/population,
         deaths_per_person = total_deaths/population) %>% 
  filter((!is.na(cases_per_person)) & (!is.na(deaths_per_person))) -> test.OCD3

test.OCD3 %>%
  group_by(date, IncomeGroup) %>%
  summarise(mean_pro_cases = mean(cases_per_person),
            mean_pro_deaths = mean(deaths_per_person)) %>%
  mutate(IncomeGroup = fct_relevel(IncomeGroup, "Low_income", after = nlevels(IncomeGroup)),
         IncomeGroup = fct_relevel(IncomeGroup, "Lower_middle_income", after = 2L)) -> INCOME_CASES_DEATHS.pro

ggplot(INCOME_CASES_DEATHS.pro, aes(date, mean_pro_cases, color = IncomeGroup)) +
  geom_line() +
  theme_bw() +
  ggtitle("Mean of cases per person")
ggplot(INCOME_CASES_DEATHS.pro, aes(date, mean_pro_deaths, color = IncomeGroup)) +
  geom_line() +
  theme_bw() +
  ggtitle("Mean deaths per person")
```

## Total cases/deaths vs date
```{r}
OCD3 %>%
  filter(!is.na(total_cases) & !is.na(total_deaths)) %>%
  group_by(date, IncomeGroup) %>%
  summarise(total_casesI = mean(total_cases),
            total_deathsI = mean(total_deaths)) %>%
  mutate(IncomeGroup = fct_relevel(IncomeGroup, "Low_income", after = nlevels(IncomeGroup)),
         IncomeGroup = fct_relevel(IncomeGroup, "Lower_middle_income", after = 2L))-> INCOME_CASES_DEATHS
  

ggplot(INCOME_CASES_DEATHS, aes(date, total_casesI, color = IncomeGroup)) +
  geom_line() +
  theme_bw() +
  ggtitle("Total cases vs date")
ggplot(INCOME_CASES_DEATHS, aes(date, total_deathsI, color = IncomeGroup)) +
  geom_line() +
  theme_bw() +
  ggtitle("Total deaths vs date")
```

## Location Taiwan 
```{r}
OCD %>%
  group_by(location, date) %>%
  summarise(total_new_cases = sum(new_cases)) -> OCD6
OCD6 %>%
  filter(location == "Taiwan") %>%
  ggplot(aes(x = date, y = total_new_cases)) +
  geom_point()
```

12. 
```{r}
OCD6 %>%
  group_by(location) %>%
  filter(!is.na(total_new_cases)) %>%
  summarise(total_new_cases = sum(total_new_cases)) %>%
  filter(location != "World" & location != "Asia" & location != "Europe" & location != "North America" & location != "North America" & location != "United States" & location != "European Union" & location != "South America") %>%
  filter(total_new_cases == max(total_new_cases) | total_new_cases == min(total_new_cases))
```

## GDP vs total cases/deaths
```{r}
OCD4 %>%
  ggplot(aes(x = gdp_per_capita, y = total_cases, color = IncomeGroup))+
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm", se = F) +
  theme_bw() +
  ggtitle("Total cases vs Catita GDP")
  
OCD4 %>%
  ggplot(aes(x = gdp_per_capita, y = total_deaths, color = IncomeGroup))+
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm", se = F) +
  theme_bw() +
  ggtitle("Total deaths vs Catita GDP")
```

## Boxplots of IncomeGroup vs cases/deaths per person
```{r}
OCD4 %>%
  ggplot(aes(x = IncomeGroup, 
             y = cases_per_person,
             )) +
  geom_boxplot(aes(fill = IncomeGroup),outlier.shape = 17, outlier.size = 3) +
  theme_bw() +
  xlab("IncomeGroup") +
  ylab("cases_per_person") +
 scale_y_log10() 

OCD4 %>%
  ggplot(aes(x = IncomeGroup, 
             y = deaths_per_person,
             )) +
  geom_boxplot(aes(fill = IncomeGroup),outlier.shape = 17, outlier.size = 3) +
  theme_bw() +
  xlab("IncomeGroup") +
  ylab("deaths_per_person") +
 scale_y_log10()
```
## Map
```{r}
library(rnaturalearth)
library(rnaturalearthdata)
library(lwgeom)
library(geos)
library(ggspatial)
library(sf)

theme_set(
  theme_void()
  )

world4<- ne_countries(scale = "medium", returnclass = "sf")
world4
full_join(world4,OCD4,by = c("adm0_a3" = "iso_code"))->newworld
class(newworld)

capitals2 <- read_csv("../data/country_capitals_long_lat.csv")

sites2 <- st_as_sf(capitals2, 
                  coords = c("Longitude", "Latitude"),
                  crs = 4326, agr = "constant")

newworld %>%
  filter(!is.na(IncomeGroup)) %>%
  group_by(IncomeGroup) %>%
  slice_max(cases_per_person, n = 2) %>%
  ungroup() %>%
  select(location, IncomeGroup, cases_per_person)-> Max5newworldc

capitals2 %>%
  filter(Country == "Seychelles" | Country == "Andorra" | Country == "Montenegro" | Country == "Georgia" | Country == "Mongolia" | Country == "Palestine" | Country == "Rwanda" | Country == "Mozambique") -> Max5capitals2c

sites2 %>%
  filter(Country == "Seychelles" | Country == "Andorra" | Country == "Montenegro" | Country == "Georgia" | Country == "Mongolia" | Country == "Palestine" | Country == "Rwanda" | Country == "Mozambique") -> Max5sites2c

ggplot(data = newworld) +
    geom_sf(aes(fill = cases_per_person)) +
    scale_fill_viridis_c(option = "plasma") +
    geom_sf(data = Max5sites2c, size = 2, shape = 23) +
    geom_text(data = Max5capitals2c, aes(x = Longitude, y = Latitude, label = Country), 
    size = 4, col = "dark blue", fontface = "bold") +
  scale_fill_continuous(low = "white", high = "blue", 
                        name = "Cases per person", label = scales::comma)
```
```{r}
newworld %>%
  filter(!is.na(IncomeGroup)) %>%
  group_by(IncomeGroup) %>%
  slice_max(deaths_per_person, n = 2) %>%
  ungroup() %>%
  select(location, IncomeGroup, deaths_per_person)-> Max5newworldd

capitals2 %>%
  filter(Country == "Hungary" | Country == "Czechia" | Country == "Peru" | Country == "Bulgaria" | Country == "Tunisia" | Country == "Ukraine" | Country == "Afghanistan" | Country == "Syria") -> Max5capitals2d

sites2 %>%
  filter(Country == "Hungary" | Country == "Czechia" | Country == "Peru" | Country == "Bulgaria" | Country == "Tunisia" | Country == "Ukraine" | Country == "Afghanistan" | Country == "Syria") -> Max5sites2d

ggplot(data = newworld) +
    geom_sf(aes(fill = deaths_per_person)) +
    scale_fill_viridis_c(option = "plasma") +
    geom_sf(data = Max5sites2d, size = 2, shape = 23) +
    geom_text(data = Max5capitals2d, aes(x = Longitude, y = Latitude, label = Country), 
    size = 4, col = "brown", fontface = "bold") +
  scale_fill_continuous(low = "white", high = "red", 
                        name = "Deaths per person", label = scales::comma)
```


## Histogram of cases/deaths per person
```{r}
ggplot(data = OCD4,aes(x = log(cases_per_person))) +
  geom_histogram()

ggplot(data = OCD4,aes(x = log(deaths_per_person))) +
  geom_histogram()

ggplot(data = OCD4,aes(x = cases_per_person)) +
  geom_histogram()+
  theme_classic()

ggplot(data = OCD4,aes(x = deaths_per_person)) +
  geom_histogram()+
  theme_classic()
```

## HDI vs cases / deaths per person
```{r}
OCD4 %>%
  filter(location != "China" & location != "Vanuatu")


ggplot(data = OCD4, aes(y = cases_per_person, x = human_development_index, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~ x,se=FALSE) +
  theme_bw()+
  ggtitle("human_development_index vs cases_per_person")

ggplot(data = OCD4, aes(y = deaths_per_person, x = human_development_index, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~ x,se=FALSE) +
  theme_bw()+
  ggtitle("Human Development Index vs Deaths per person")
```
## Population vs cases / deaths per person
```{r}
OCD4 %>%
  filter(location != "China" & location != "India" & location != "United States") %>%
ggplot(aes(y = (cases_per_person), x = (population), color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm",se=FALSE)

OCD4 %>%
  filter(location != "China" & location != "India" & location != "United States") %>%
ggplot(aes(y = (deaths_per_person), x = (population), color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm",se=FALSE)

OCD4 %>%
  group_by(IncomeGroup) %>%
  slice_max(population)
```

## Median age vs cases / deaths per person
```{r}
OCD4 %>% 
filter(location!="China") %>% 
  ggplot(aes(y = cases_per_person, x = median_age, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm",se=FALSE)+
  scale_y_log10()+
  theme_bw()

OCD4 %>% 
filter(location!="China") %>% 
  ggplot(aes(y = deaths_per_person, x = median_age, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm",se=FALSE)+
  scale_y_log10()+
  theme_bw()
```

## Population density vs cases / deaths per person
```{r}
ggplot(data = OCD4,aes(y = (cases_per_person), x = (population_density), color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm")+
  scale_x_log10()+
  theme_bw()

ggplot(data = OCD4,aes(y = (deaths_per_person), x = population_density, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm")+
  scale_x_log10()+
  theme_bw()
```

## Diabetes vs cases / deaths per person
```{r}
OCD4 %>%
  filter(location != "Mauritius" & location != "Saudi Arabia") %>%
ggplot(aes(y = (cases_per_person), x = diabetes_prevalence, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm")

OCD4 %>%
  filter(location != "Mauritius" & location != "Saudi Arabia") %>%
ggplot(aes(y = (deaths_per_person), x = diabetes_prevalence, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm")

OCD4 %>%
  #filter(location != "Mauritius" & location != "Saudi Arabia") %>%
ggplot(aes(y = (cases_per_person), x = diabetes_prevalence, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm")+
  scale_y_log10()+
  theme_bw()

OCD4 %>%
  #filter(location != "Mauritius" & location != "Saudi Arabia") %>%
ggplot(aes(y = (deaths_per_person), x = diabetes_prevalence, color = IncomeGroup)) +
  geom_point() +
  geom_smooth(method = "lm")+
  scale_y_log10()+
  theme_bw()
```

## Fit the model
```{r}
OCD4.lm1 <- lm(log(cases_per_person) ~ human_development_index + IncomeGroup + log(gdp_per_capita) + median_age +log(population_density), data = Test.OCD4)

OCD4.lm2 <- lm(log(deaths_per_person) ~ human_development_index + IncomeGroup + log(gdp_per_capita) + median_age +log(population_density), data = Test.OCD4)

summary(OCD4.lm1)
summary(OCD4.lm2)
```

## Check the constance
```{r}
par(mfrow=c(3,2))

plot(Test.OCD4$median_age, OCD4.lm1$residuals)
abline(0, 0, lty=2)

plot(log(Test.OCD4$gdp_per_capita),OCD4.lm1$residuals)
abline(0, 0, lty=2)

plot(log(Test.OCD4$population_density), OCD4.lm1$residuals)
abline(0, 0, lty=2)

plot(Test.OCD4$human_development_index, OCD4.lm1$residuals)
abline(0, 0, lty=2)

plot(Test.OCD4$IncomeGroup, OCD4.lm1$residuals)
abline(0, 0, lty=2)

par(mfrow=c(3,2))
plot(Test.OCD4$median_age, OCD4.lm2$residuals)
abline(0, 0, lty=2)

plot(log(Test.OCD4$gdp_per_capita),OCD4.lm2$residuals)
abline(0, 0, lty=2)

plot(log(Test.OCD4$population_density), OCD4.lm2$residuals)
abline(0, 0, lty=2)

plot(Test.OCD4$human_development_index, OCD4.lm2$residuals)
abline(0, 0, lty=2)

plot(Test.OCD4$IncomeGroup, OCD4.lm2$residuals)
abline(0, 0, lty=2)
```
## Check normality of residual
```{r}
hist(OCD4.lm1$resi)
hist(OCD4.lm2$resi)
```

## Stepwise
```{r}
library(olsrr)
OCD4.lm1.step <- ols_step_both_aic(OCD4.lm1)
OCD4.lm1.step
```

```{r}
OCD4.lm2.step <- ols_step_both_aic(OCD4.lm2)
OCD4.lm2.step
```


## Final model
```{r}
Final.model1 <- lm(log(cases_per_person) ~ human_development_index + IncomeGroup, data = Test.OCD4)
summary(Final.model1)
```

```{r}
Final.model2 <- lm(log(deaths_per_person) ~ human_development_index + IncomeGroup + log(gdp_per_capita), data = Test.OCD4)
summary(Final.model2)
```

## Check the normality of the variable
```{r}
ggplot(data = Final.model1, aes(x = Final.model1$fitted.values, y = Final.model1$residuals, color = IncomeGroup)) +
  geom_point() +
  geom_hline(yintercept=0,linetype="dashed") +
  ggtitle("Model residuals vs model fitted values (Cases)")

qqnorm(Final.model1$resi)
qqline(Final.model1$resi)
shapiro.test(Final.model1$resi)
```
## Diagnostics
```{r}
#Cases
ggplot(data = Final.model1, aes(x = Final.model1$fitted.values, y = Final.model1$residuals, color = IncomeGroup)) +
  geom_point() +
  geom_hline(yintercept=0,linetype="dashed") +
  ggtitle("Model residuals vs model fitted values (Cases)")

qqnorm(Final.model1$resi)
qqline(Final.model1$resi)
shapiro.test(Final.model1$resi)

# Deaths
ggplot(data = Final.model2, aes(x = Final.model2$fitted.values, y = Final.model2$residuals, color = IncomeGroup)) +
  geom_point() +
  geom_hline(yintercept=0,linetype="dashed") +
  ggtitle("Model residuals vs model fitted values (Deaths)")

qqnorm(Final.model2$resi)
qqline(Final.model2$resi)
shapiro.test(Final.model2$resi)
```
## Residuals vs preorder Resisual
```{r}
ggplot(data = Test.OCD4, aes(x = lag(Final.model1$residuals), y = Final.model1$residuals)) +
  geom_point() +
  geom_hline(yintercept=0,linetype="dashed") +
  ggtitle("Residuals vs preorder Residual (Cases)")

ggplot(data = Test.OCD4, aes(x = lag(Final.model2$residuals), y = Final.model2$residuals)) +
  geom_point() +
  geom_hline(yintercept=0,linetype="dashed") +
  ggtitle("Residuals vs preorder Residual (Deaths)")
```

## Check multicollinearity
```{r}
library(car)
vif(Final.model1)
vif(Final.model2)
```


